### 2021-10-21 proofs/inversion-of-typing.tex

\begin{lemma}[Inversion of typing]
  \lbl{lem:inversion-of-typing}~
  \begin{enumerate}
    \item % 1
    If $\Γ \⊢ \λ \x \: \S_{1}.~ \s_{2} \: \T$ and $\Γ \⊢ \T \<: \U_{1} \→ \U_{2}$,
    then $\Γ \⊢ \U_{1} \<: \S_{1}$ and there is some $\S_{2}$ such that
    $\Γ, \x \: \S_{1} \⊢ \s_{2} \: \S_{2}$ and $\Γ \⊢ \S_{2} \<: \U_{2}$.

    \item % 2
    If $\Γ \⊢ \λ \X \<: \S_{1}.~ \s_{2} \: \T$ and $\Γ \⊢ \T \<: \(\∀ \X \<: \U_{1}.~ \U_{2})$,
    then $\U_{1} \= \S_{1}$ and there is some $\S_{2}$ such that
    $\Γ, \X \<: \S_{1} \⊢ \s_{2} \: \S_{2}$ and $\Γ, \X \<: \S_{1} \⊢ \S_{2} \<: \U_{2}$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  \begin{enumerate}
    \item % 1
    By induction on a derivation of $\Γ \⊢ \t \: \T$.

    \begin{itemize}
      \Case\TAbs:
      \quad $\t \= \λ \x \: \S_{1} \s_{2}$
      \quad $\T \= \S_{1} \→ \T_{2}$
      \quad $\Γ \, \x \: \S_{1} \⊢ \s_{2} \: \T_{2}$

      Given $\Γ \⊢ \S_{1} \→ \T_{2} \<: \U_{1} \→ \U_{2}$,
      we use \Cref{lem:inversion-of-subtyping} to obtain
      $\Γ \⊢ \U_{1} \<: \S_{1}$ and $\Γ \⊢ \T_{2} \<: \U_{2}$.
      We pick $\S_{2}$ to be $\T_{2}$ to obtain the desired result.

      \Case\TSub:
      \quad $\Γ \⊢ \t \: \S$
      \quad $\Γ \⊢ \S \<: \T$

      Using \STrans with $\Γ \⊢ \S \<: \T$ and $\Γ \⊢ \T \<: \U_{1} \→ \U_{2}$ we get $\Γ \⊢ \S \<: \U_{1} \→ \U_{2}$. The results follows directly from the IH.

      \Case\TVar, \TApp, \TTAbs, \TTApp, \TClass, \TMatch:

      In those cases, $\t$ is not of the form $\λ \x \: \S_{1}.~ \s_{2}$ and the result is immediate.
    \end{itemize}

    \item % 2
    By induction on a derivation of $\Γ \⊢ \t \: \T$.

    \begin{itemize}
      \Case\TTAbs:
      \quad $\t \= \λ \X \<: \S_{1}.~ \s_{2}$
      \quad $\T \= \∀ \X \<: \S_{1}.~ \T_{2}$
      \quad $\Γ, \X \<: \S_{1} \⊢ \s_{2} \: \T_{2}$

      Given $\Γ \⊢ \(\∀ \X \<: \S_{1}.~ \T_{2}) \<: \(\∀ \X \<: \U_{1}.~ \U_{2})$,
      we use \Cref{lem:inversion-of-subtyping} to obtain
      $\U_{1} \= \S_{1}$ and $\Γ, \X \<: \S_{1} \⊢ \T_{2} \<: \U_{2}$.
      We pick $\S_{2}$ to be $\T_{2}$ to obtain the desired result.

      \Case\TSub:
      \quad $\Γ \⊢ \t \: \S$
      \quad $\Γ \⊢ \S \<: \T$

      Using \STrans with $\Γ \⊢ \S \<: \T$ and $\Γ \⊢ \T \<: \(\∀ \X \<: \U_{1}.~ \U_{2})$
      we get $\Γ \⊢ \S \<: \(\∀ \X \<: \U_{1}.~ \U_{2})$.
      The results follows directly from the IH.

      \Case\TVar, \TAbs, \TApp, \TTApp, \TClass, \TMatch:

      In those cases, $\t$ is not of the form $\λ \X \<: \S_{1}.~ \s_{2} \: \T$ and the result is immediate.

    \end{itemize}
  \end{enumerate}
\end{proof}