### 2021-10-21 proofs/minimum-types.tex

\begin{lemma}[Minimum types]
  \lbl{lem:minimum-types}~
  \begin{enumerate}
    \item % 1
    If $\Γ \⊢ \new \C \: \T$ then $\Γ \⊢ \{ \new \C \} \<: \T$.

    \item % 2
    If $\Γ \⊢ \λ \x \: \T_{1}.~ \t_{2} \: \T$
    then there is some $\T_{2}$ such that
    $\Γ \⊢ \T_{1} \→ \T_{2} \<: \T$.

    \item % 3
    If $\Γ \⊢ \λ \X \<: \U_{1}.~ \t_{2} \: \T$
    then there is some $\T_{2}$ such that $\Γ \⊢ \∀ \X \<: \U_{1}.~ \T_{2} \<: \T$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  \begin{enumerate}
    \item % 1
    By induction on a derivation of $\Γ \⊢ \t \: \T$.
    \begin{itemize}
      \Case\TVar, \TAbs, \TApp, \TTAbs, \TTApp, \TMatch:

      In those cases, $\t$ is not a constructor call and the result is immediate.

      \Case\TClass:
      \quad $\t \= \new \C$
      \quad $\T \= \{ \new \C \}$

      The result follows directly from \SRefl.

      \Case\TSub:
      \quad $\Γ \⊢ \t \: \S$
      \quad $\Γ \⊢ \S \<: \T$

      By the IH, $\Γ \⊢ \{ \new \C \} \<: \S$.
      The result follows from \STrans.
    \end{itemize}

    \item % 2
    By induction on a derivation of $\Γ \⊢ \t \: \T$.
    \begin{itemize}
      \Case\TVar, \TApp, \TTAbs, \TTApp, \TClass, \TMatch:

      In those cases, $\t$ is not an abstraction and the result is immediate

      \Case\TAbs:
      \quad $\T \= \T_{1} \→ \S_{2}$
      \quad $\Γ, \x \: \T_{1} \⊢ \t_{2} \: \S_{2}$

      $\T_{2} \= \S_{2}$ and the result is immediate using \SRefl.

      \Case\TSub:
      \quad $\Γ \⊢ \t \: \S$
      \quad $\Γ \⊢ \S \<: \T$

      Using the IH, there is some $\T_{2}$ such that $\Γ \⊢ \T_{1} \→ \T_{2} \<: \S$.
      The result follows from \STrans.
    \end{itemize}

    \item % 3
    By induction on a derivation of $\Γ \⊢ \t \: \T$.
    \begin{itemize}
      \Case\TVar, \TAbs, \TApp, \TTApp, \TClass, \TMatch:

      In those cases, $\t$ is not a type abstraction and the result is immediate

      \Case\TTAbs:
      \quad $\T \= \∀ \X \<: \U_{1}.~ \S_{2}$
      \quad $\Γ, \X \<: \U_{1} \⊢ \t_{2} \: \S_{2}$

      $\T_{2} \= \S_{2}$ and the result is immediate using \SRefl.

      \Case\TSub:
      \quad $\Γ \⊢ \t \: \S$
      \quad $\Γ \⊢ \S \<: \T$

      Using the IH, there is some $\T_{2}$ such that $\Γ \⊢ \∀ \X \<: \U_{1}.~ \T_{2} \<: \S$.
      The result follows from \STrans.
    \end{itemize}
  \end{enumerate}
\end{proof}