### 2021-10-22 code-sections.tex
Import TLPwMT

\newcommand{\goodreducescalacodesection}{\lstinputlisting[style=scala, firstline=3, lastline=4, breaklines]{scala/src/main/scala/goodreduce.scala}}
\newcommand{\shapecodesection}{\lstinputlisting[style=scala, firstline=8, lastline=11, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\datatypecodesection}{\lstinputlisting[style=scala, firstline=19, lastline=19, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\ndarraycodesection}{\lstinputlisting[style=scala, firstline=23, lastline=23, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\randomnormalcodesection}{\lstinputlisting[style=scala, firstline=27, lastline=27, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\multiplycodesection}{\lstinputlisting[style=scala, firstline=31, lastline=31, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\numelementscodesection}{\lstinputlisting[style=scala, firstline=36, lastline=40, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\reshapecodesection}{\lstinputlisting[style=scala, firstline=44, lastline=45, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\reducecodesection}{\lstinputlisting[style=scala, firstline=50, lastline=66, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\npmeancodesection}{\lstinputlisting[style=scala, firstline=86, lastline=87, breaklines]{scala/src/main/scala/numpy.scala}}
\newcommand{\badreducescalacodesection}{\lstinputlisting[style=scala, firstline=2, lastline=4, breaklines]{scala/src/test/scala/badreduce.scala}}