### 2022-02-14 ack.tex
Acknowledgements

Advisor:

Martin Odersky

Thesis Committee:

Rachid Guerraoui
Paolo Giarrusso
Richard Eisenberg
Viktor Kuncak

Scala center colleagues:

Guillaume Masse
Jorge Vicente
Ólafur Páll Geirsson
Travis Lee
Darja Jovanovic
Jamie Thompson
Julien Richard-Foy

LAMP colleagues:

Aggelos Biboudis
Aleksander Boruch-Gruszecki
Allan Renucci
Anatolii Kmetiuk
Denys Shabalin
Dmitry Petrashko
Felix Mulder
Fengyun Liu
Jonathan Brachthäuser
Manohar Jonnalagedda
Matt Bovel
Nicolas Alexander Stucki
Paolo Giarrusso
Sébastien Doeraene
Sandro Stucki

LAMP colleagues:

Georg Schmid
Jad Hamza

Lightbend:

Adriaan Moors
Dale Wijnand

Students:

Maxime Kjaer
Andrea Veneziano

Staff:

Fabien Salvi
Natascha Fontana
Faille France

Friends:

Mia Primorac
Hermina Petric Maretić

Family:

Alicia Blanvillain
Soledad Andany
Christian Blanvillain
Leila

Girlfriend:

Dragana Milovancevic