### 2022-02-14 ack.tex
Acknowledgements

First and foremost, I would like to thank my advisor, Martin Odersky, for his help and support throughout my PhD study, and for providing me with this unique opportunity to work on my favorite language, Scala.

I would also like to thank the other members of my thesis committee, Paolo Giarrusso, Richard Eisenberg, and Viktor Kuncak, for agreeing to review my dissertation and taking part in the committee.

I would like to thank all my colleagues from LAMP, LARA and Scala Center.
You were the source of so many interesting discussions, cheerful moments, and great memories, it was really a pleasure to work along your side.
Aggelos Biboudis,
Aleksander Boruch-Gruszecki,
Allan Renucci,
Anatolii Kmetiuk,
Darja Jovanovic,
Denys Shabalin,
Dmitry Petrashko,
Fabien Salvi,
Felix Mulder,
Fengyun Liu,
Guillaume Martres,
Guillaume Masse,
Jad Hamza,
Jamie Thompson,
Jonathan Brachthäuser,
Jorge Vicente,
Julien Richard-Foy,
Manohar Jonnalagedda,
Matt Bovel,
Natascha Fontana,
Nicolas Alexander Stucki,
Ólafur Páll Geirsson,
Paolo Giarrusso,
Romain Edelmann,
Sandro Stucki,
Sébastien Doeraene,
Simon Guilloud, and
Travis Lee---thank you!
If it wasn't for you, I would have not have gotten here.

A special thanks to Georg Schmid for being such an amazing friend, colleague and co-author over the past five years.
I will never forget what we went through together, the paper deadlines, the hacking sessions, the bike rides, all games...
It's crazy how much we did and how much we learned together.
And of course, thank you Mia Primorac for being there for all the fun parts!

I would like to thank my parents, Soledad and Christian, and my little sister, Alicia, for being there for me all these years.
It was truly a pleasure to have you close by throughout my PhD.
I must also thank the latest arrival in family, Leila, for being such an endless source of joy and cuteness.

Last but not the least, I would like to thank Dragana Milovancevic for being always there on my side, for being so kind and patient, and for her tremendous help in putting the ideas of this thesis into text.
You are the best.